---
title: "mcd_analysis"
author: "carmen zheng"
date: "`r Sys.Date()`"
output: 
  html_document:
    theme: united
    toc: yes
    toc_float:
      collapsed: true
---

This is the master script for **STUDY 2 - analysis of TeleNP data for MCD patients in AEP**

```{r}
rm(list=ls())
```

# Libraries  

```{r}
library(tidyverse)
library(ggplot2)
library(pwr) # power analysis
```

# Load data

```{r}
# the patient characteristics data
mcd_data = read.csv("/Users/carmen/Library/CloudStorage/OneDrive-TheUniversityofMelbourne/PhD/data/study-2/AustralianEpilepsyPr-MCDPatientCharacteri_DATA_2024-05-14_1100.csv")

# the TeleNP data
telenp_data <- load("/Users/carmen/Library/CloudStorage/OneDrive-TheUniversityofMelbourne/PhD/data/study-2/allPtntCtrlData.RData")

# the REDCap npsych data of all MCD patients
redcap_mcd_data <- read.csv("/Users/carmen/Library/CloudStorage/OneDrive-TheUniversityofMelbourne/PhD/data/study-2/AustralianEpilepsyPr-MCDPatientsNpsychSco_DATA_2024-05-14_1102.csv")
```

# Data cleaning

## For patient characeristics data:

### Remove meaningless columns

```{r}
mcd_data <- mcd_data[-c(2:4)] # remove redcap generated event indices
```

### Rename columns

```{r}
mcd_data <- mcd_data %>%
  rename(
    pt_sex = aepref_pt_sex,
    pt_age = npsyts_age_at_npsych_ax,
    pt_ref = aepref_pt_ref_cat, 
    # imaging outcome classification status:
    pt_abnorm_status = mridat_rad_rep_sum_class,
    pt_ep_lesn_status = mridat_epileptogenic_lesn, 
    pt_motion_status = mridat_mri_rad_motion_aff,
    pt_repeat_status = mridat_mri_rad_more_img,
    # MCD lesion types:
    MCD_BOSD = mridat_rad_class_mcd___1, # BOSD
    MCD_FCD = mridat_rad_class_mcd___2, # FCD, not BOSD
    MCD_tuber = mridat_rad_class_mcd___3, # tuberous sclerosis
    MCD_GMH_PVNH = mridat_rad_class_mcd___4, # grey matter heterotopia - PVNH
    MCD_GMH_subnod = mridat_rad_class_mcd___5, # grey matter heterotopia - subcortical nodular
    MCD_GMH_band = mridat_rad_class_mcd___6, # grey matter heterotopia - band heterotopia
    MCD_polymicrogyria = mridat_rad_class_mcd___7, # polymicrogyria
    MCD_hh = mridat_rad_class_mcd___8, # hypothalamic hamartoma
    MCD_meningocele = mridat_rad_class_mcd___9, # meningocele
    MCD_encephalocele = mridat_rad_class_mcd___10, # encephalocele
    MCD_CC_dysgenesis = mridat_rad_class_mcd___11, # corpus callosum - agenesis/dysgenesis
    MCD_chiari = mridat_rad_class_mcd___12, # chiari malformation
    MCD_other = mridat_rad_class_mcd___13, # other (schizencephaly, lissencephaly, hemimegalencephaly etc)
    MCD_none = mridat_rad_class_mcd___14, # none
    MCD_unknown = mridat_rad_class_mcd___15 # unknown
  )

```

### Make categorical variables factors

```{r echo=TRUE}
# sex
mcd_data$pt_sex <- factor(mcd_data$pt_sex,
                      levels = c(1,2,3),
                      labels = c("M", "F", "O"))

# referral category
mcd_data$pt_ref <- factor(mcd_data$pt_ref,
                      levels = c(1,2,3),
                      labels = c("FUS", "New.Dx", "DRE"))

# imaging status - abnormality
mcd_data$pt_abnorm_status <- factor(mcd_data$pt_abnorm_status,
                                levels = c(1,2,3),
                                labels = c("normal", "abnormal", "incidental"))

# imaging status - epileptogenic lesion
mcd_data$pt_ep_lesn_status <- factor(mcd_data$pt_ep_lesn_status,
                                 levels = c(1,2,3),
                                 labels = c("yes", "no", "uncertain"))

# imaging status - motion
mcd_data$pt_motion_status <- factor(mcd_data$pt_motion_status,
                                levels = c(1,2,3),
                                labels = c("not.mentioned", "mild", "degraded"))

# imaging status - repeat scan recommended
mcd_data$pt_repeat_status <- factor(mcd_data$pt_repeat_status,
                                levels = c(1,2),
                                labels = c("yes", "no"))

head(mcd_data)
```

### Patient counts

```{r}
# create a variable containing all unique id
unique_id = unique(mcd_data$record_id) 

# create variable containing all MCD columns
mcd_types = c("MCD_BOSD",
              "MCD_FCD",
              "MCD_tuber",
              "MCD_GMH_PVNH",
              "MCD_GMH_subnod",
              "MCD_GMH_band",
              "MCD_polymicrogyria",
              "MCD_hh",
              "MCD_meningocele",
              "MCD_encephalocele",
              "MCD_CC_dysgenesis",
              "MCD_chiari",
              "MCD_other") 

# to identify pts with varying numbers of MCD, make a new column to store lesion count for each
mcd_data$num_mcd = NaN 

# for each id, if the sum of lesion is six/five/four/three/two/one/zero, code this numerical sum in colname = num_mcd
for (id in unique_id) {
  # pull the data for this record
  t.record = which(mcd_data$record_id == id)
  # get MCD columns for this patient
  t.mcd.data = mcd_data[t.record, mcd_types] 
  if (rowSums(t.mcd.data, na.rm = TRUE) == 6) {
    # there are exactly six MCDs 
    mcd_data[t.record, "num_mcd"] = 6
  }
  else if (rowSums(t.mcd.data, na.rm = TRUE) == 5) {
    # there are exactly five MCDs 
    mcd_data[t.record, "num_mcd"] = 5
  }
  else if (rowSums(t.mcd.data, na.rm = TRUE) == 4) {
    # there are exactly four MCDs 
    mcd_data[t.record, "num_mcd"] = 4
  } else if (rowSums(t.mcd.data, na.rm = TRUE) == 3) {
    # there are exactly three MCDs 
    mcd_data[t.record, "num_mcd"] = 3
  } else if (rowSums(t.mcd.data, na.rm = TRUE) == 2) {
    # there are exactly two MCDs 
    mcd_data[t.record, "num_mcd"] = 2
  } else if (rowSums(t.mcd.data, na.rm = TRUE) == 1) {
    # there is exactly one MCD 
    mcd_data[t.record, "num_mcd"] = 1
  } else {
    # no relevant MCD identified
    mcd_data[t.record, "num_mcd"] = 0
  }
} 

num_mcd <- factor(mcd_data$num_mcd, 
                  levels = c(0, 1, 2, 3, 4, 5, 6))
table(num_mcd)

# count number of cases per each MCD in single-pathology cases
one_path <- subset(mcd_data, mcd_data$num_mcd == 1)
one_path %>% summarise(across(mcd_types, sum))

# count number of cases per each MCD in dual-pathology cases
two_path <- subset(mcd_data, mcd_data$num_mcd == 2)
two_path %>% summarise(across(mcd_types, sum)) 

# count number of cases per each MCD in triple-pathology cases
three_path <- subset(mcd_data, mcd_data$num_mcd == 3)
three_path %>% summarise(across(mcd_types, sum)) 

# count number of cases per each MCD in quadruple-pathology cases
four_path <- subset(mcd_data, mcd_data$num_mcd == 4)
four_path %>% summarise(across(mcd_types, sum)) 

# count number of cases per each MCD in quintuple-pathology cases
five_path <- subset(mcd_data, mcd_data$num_mcd == 5)
five_path %>% summarise(across(mcd_types, sum)) 
```

## for REDCap data:

### Remove meaningless columns

```{r}
redcap_mcd_data <- redcap_mcd_data[-c(2:4)] # removed redcap generated event indices
```


### Rename columns

```{r}
# match columns common for both redcap_mcd_data and telenp_data
redcap_mcd_data <- redcap_mcd_data %>% rename(
  gender = aepref_pt_sex,
  ax.age.rough = npsyts_age_at_npsych_ax,
  # RAVLT / word-list learning
  wll.trial1.total = npsyts_ravlt_t1_raw_sc,
  wll.trial2.total = npsyts_ravlt_t2_raw_sc,
  wll.trial3.total = npsyts_ravlt_t3_raw_sc,
  wll.total = npsyts_ravlt_t1_3_autocal,
  wld.total = npsyts_ravlt_delay_raw_sc,
  
  # digit span
  dsp.ldsf = npsyts_dsb_ldsf,
  dsp.ldsb = npsyts_dsb_ldsb, 
  
  # oral SDMT
  sdm.total.correct = npsyts_oral_sdmt_raw, 
  
  # COWAT / BHR
  lf.B.valid = npsyts_cowat_b,
  lf.H.valid = npsyts_cowat_h,
  lf.R.valid = npsyts_cowat_r,
  lf.total.valid = npsyts_bhr_total_raw,
  
  # animals / category fluency
  cf.total.valid = npsyts_animals_raw_score,
  
  # confrontation naming
  cnt.total.correct = npsyts_cnt_total
)
```

### Make categorical variables factors
```{r}
# re-define gender in redcap_mcd_data
redcap_mcd_data$gender <- factor(redcap_mcd_data$gender,
                                 levels = c(1,2,3),
                                 labels = c("Male", "Female", "Other"))
```

# Merge REDCap and TeleNP data

1. a list of MCD patients who are on TeleNP --- `telenp_mcd_pts`
2. TeleNP scores of only MCD patients --- `telenp_mcd_data`
3. npsych scores of only MCD patients not on TeleNP --- `redcap_mcd_data`
    - hence rowsum of `telenp_mcd_data` and `redcap_mcd_data` = total sample size
4. initialise new df to store `redcap_mcd_data` and `telenp_mcd_data` --- `cog_mcd_data`


```{r}
# find those record id of mcd patients who are on TeleNP
telenp_mcd_pts <- intersect(mcd_data$record_id, ptnt.df$record_id)

# make telenp_mcd_data contain npsych scores of only those on TeleNP
telenp_mcd_data <- data.frame() # initialise data frame
telenp_mcd_data <- ptnt.df[(ptnt.df$record_id %in% telenp_mcd_pts), ]

# make record_id the first col in telenp_mcd_data
telenp_mcd_data <- telenp_mcd_data %>% relocate(record_id)

# make redcap_mcd_data contain npsych scores of only those not on TeleNP 
redcap_mcd_data <- redcap_mcd_data[!(redcap_mcd_data$record_id %in% telenp_mcd_pts), ]

# initialise a new df for analysis
cog_mcd_data <- data.frame()
cog_mcd_data <- bind_rows(telenp_mcd_data, redcap_mcd_data)
```

# Analysis
## Power analysis
```

# Troubleshoot

- [x] Check what's up with the two people with no MCD lesions 
  - it was because i didn't specify what happens when people have more than three MCDs